---
title: "Visualizing Protein Structure"
author: "Vikas Gubbala"
date: "2/13/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Structural Analysis of HIV-1 Protease

HIV-1 protease is essential for the formation of HIV proteins. Inhibition of this protein may aid in the suppress viral replication. In this analysis, we will look at the PDB file 1HSG, which contains the protease bound to the drug indinavir.

PDB files contain lists of atoms along with coordinates. Sometimes, additional information like secondary structure is added.

```{r}
library(bio3d)
library(ggplot2)
library(ggrepel)
```

## Working with Single PDB files.

The 1HSG file is read from the online database.

```{r}
#accesses online pdb file
pdb <- read.pdb("1hsg")

# summary of pdb object
pdb 
```

We can select for certain attributes as well.

`attributes(pdb)` - returns attributes of the object
`pdb$atom` - selection of the attribute 'atom'
`pdb$atom[rows, cols]` - subset of atoms


```{r}
#attributes of PDB object
attributes(pdb) 

#selects an attribute of the object
head(pdb$atom) 

#FINISH THIS

protein <- atom.select(pdb, "protein", value = TRUE)
ligand <- atom.select(pdb, "ligand", value = TRUE)
```



Protein Visualization is done in the software VMD.

```{r}
## Do this!!
```

## Aligning multiple PDB files

For this, MUSCLE is needed.