---
title: "Analysis of Human Breast Cancer Data"
author: "Vikas Gubbala"
date: "2/6/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing the data into R
```{r}
library(ggplot2)

fna.data <- "data/WisconsinCancer.csv"
wisc.df <- read.csv("data/WisconsinCancer.csv")

new <- read.csv("data/new_samples.csv")
```

Preparing the Data
```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
row.names(wisc.data) <- wisc.df$id
diagnosis <- wisc.df[,2]

head(wisc.data)
#head(diagnosis)
```

Exploratory Data Analysis
```{r}
print(paste("Number of rows = ", nrow(wisc.data)))
print(paste("Number of malignant samples = ", sum(diagnosis == "M")))
print(paste("Number of avg characteristics = ", sum(grepl("_mean", colnames(wisc.data)))))
```


Performing PCA
```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
#summary(wisc.pr)
```
From the results it seems that 99% of variation can be explained by PC1 captures 44% of the variation in the original data set, and PC2 captures 19%. Three PCs are required to capture 70% of the variance of the original dataset, and seven PCs are required to capture 90% of the variance. 


Interpreting PCA results
```{r}
pr_df <- as.data.frame(wisc.pr$x)
pr_df$group <- diagnosis
plot <- ggplot(pr_df,aes(x=PC1,y=PC2,color=group)) + geom_point(pch = 1) + theme_classic() + scale_colour_discrete(name = "Diagnosis", breaks = c("B", "M"), labels = c("Benign", "Malignant"))
plot
```

Explaining Variance
```{r}
pr.var = wisc.pr$sdev^2
pve <- pr.var / sum(pr.var)

#pve_plot <- ggplot(as.data.frame(pve)) + geom_bar()
#pve_plot

barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

Communicating PCA Results

```{r}
#finish this part!
```

Clustering in PC Space

```{r}
#Dendrogram Clustering
wisc.pr.hclust = hclust(dist(wisc.pr$x[,1:3]), method = "ward.D2")
plot(wisc.pr.hclust, labels = FALSE)
groups <- cutree(wisc.pr.hclust, k=2)
table(groups, diagnosis)
```

```{r}
#Graphing PCA by dendrogram clustering
wisc.pr.hclust = hclust(dist(wisc.pr$x[,1:3]), method = "ward.D2")
groups <- cutree(wisc.pr.hclust, k=2)
table(groups, diagnosis)
plot <- ggplot(pr_df,aes(x=PC1,y=PC2)) + geom_point(pch = 1) + theme_classic() + scale_colour_discrete(name = "Diagnosis", breaks = c("B", "M"), labels = c("Benign", "Malignant"))
g <- as.factor(3-groups)
plot + aes(color = g)
```

Specificity Versus Sensitivity

Minimizing FPs and FNs by altering clustering method, linkage method.

Prediction

```{r}
npc <- predict(wisc.pr, newdata=new)
npc
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```


